A comparison of different methods for calculating tangent-stiffness matrices in a massively parallel computational peridynamics code.

### 2014-02-27 response/response.tex

{\color{red}  
\begin{itemize}
     \item
      The word, 'discrimination' has been removed from the manuscript. 'Discrimination' was meant to mean 'measuring differences and making judgements on relative usefulness'. Since this meaning is probably understood from the word 'comparison' alone, the word 'discrimination' has been removed from the manuscript for clarity.
  \end{itemize}}

` \item	
	Pg. 2: In line 37, explain what do you mean by “in-situ instrumentation”.

{\color{red}  
\begin{itemize}
     \item
      The words, 'in-situ' have been removed from the manuscript. 'In-situ' was meant to suggest that timing measurements were taken from within and by the program being analyzed itself, rather than by an external program, but it has been removed for clarity.
  \end{itemize}}

  \item 
	Pg. 3: In lines 19-29, specify the sections where each part of the manuscript is
included.

{\color{red}  
\begin{itemize}
     \item
      The correction is made in the manuscript.
  \end{itemize}}

  \item 
Pg. 3: You may consider including the exact web address of the paper repository.

{\color{red}  
\begin{itemize}
     \item
      The correction is made in the manuscript. An exact web address for the paper repository has been added.
  \end{itemize}}

  \item 
Pg. 3: Please include the forward and centered ﬁnite difference expressions for
easier comparison with (2); include the truncation error as well. You do not have to
derive these expressions though.

{\color{red}  
\begin{itemize}
     \item
      The correction is made in the manuscript. Expressions for forward and centered difference have been included for faster comparison with (2).  
  \end{itemize}}

  \item 
Pg. 4: In Eq. (1), “$\partial f^2$” should be “$\frac{\partial^2 f}{\partial x^2}$”; similarly, for the third derivative.

{\color{red}  
\begin{itemize}
     \item
      The correction is made in the manuscript.
  \end{itemize}}

  \item 
Pg. 4: You may consider writing Eq. (2) with the truncation error

{\color{red}  
\begin{itemize}
     \item
      The correction is made in the manuscript.
  \end{itemize}}

9. Pg. 6: In lines 9-11, please add references to the “Poisson problem” and to the
“mathematical optimization”.

10. Pg. 7: Should Eqs. (4), (5), and (6) use “$\approx$” instead of “$=$”?

11. Pg. 7: In line 28, note that (cid:126)u is not the current conﬁguration, but the displacement
ﬁeld. Please revise the sentence.

12. Pg. 7: Please provide an expression for ${F^int}_i$
in the model used.

13. Pg. 9: In line 26, please add the reference [12] below to “coupling to molecular
simulations [29].”

14. Pg. 10: Please add an illustration of the block undergoing tension in Section 2.2,
for clarity. Also, clarify the displacement boundary conditions; are they applied in a
volumetric layer?

15. Pg. 10: In line 40, you mentioned a bandwidth of 7. This seems to be clear in 1D.
Does this hold in higher dimensions? Please clarify.

16. Pg. 10: In lines 47-49, you say “The speciﬁc parameters ... can be found ... in the
paper repository.” It seems a good idea to have the paper repository with ﬁles that a
user can eventually use. However, the content of the paper should be independent
of the repository. Please include a table with the parameters used. Similarly, in
page 11 (lines 11-13).

17. Pg. 14: In line 55, is there any conclusion you can draw from the power law index
$1.0 * 10^-6$?
{\color{red}  
\begin{itemize}
     \item
      Manuscript updated. More significant than the power law index is that the methods have the same power law relationship with computation time at all, meaning that there is some, but weak evidence that the methods have equal algorithm complexity. 
  \end{itemize}}

18. Pgs. 14-: Please clarify the value of the meshsize used on each of the experiments.
{\color{red}  
\begin{itemize}
     \item
      [TODO]
  \end{itemize}}

19. Pg. 16: Would it be possible to rescale Fig. 2, so that the dependence of the time
on the number of cores will look indeed linear?
{\color{red}  
\begin{itemize}
     \item
     Manuscript is in error and has been corrected. An accurate statement is that CPU time and number of cores are inversely related.
  \end{itemize}}

20. Pg. 16: In lines 54-56, you say “It should be pointed out that compared to the (cid:96)2 -
norms of any of the TS matrices by themselves, the magnitude of the accuracy
metric ... is relatively small.” Please include in the manuscript the explicit informa-
tion on the relative error. It seems that you already have such information in the
paper repository.
{\color{red}  
\begin{itemize}
     \item
[TODO]
  \end{itemize}}

21. Pg. 17: Include the values of the probe distances h used for each method.
{\color{red}  
\begin{itemize}
     \item
[TODO]
     The probe distances used in Peridigm are automatically determined with a heuristic algorithm and are not set by the user. They are 1.E-6 times the node spacing [ADAGIO REF] for FD and CD which is determined to limit the onset of subtractive cancellation. 1.E-100 was hard-coded for CS, while AD had no probe distance.
  \end{itemize}}

22. Pg. 17: In lines 53-54, you say “Accuracy trends matched those for the single core
tests.” Indeed, comparing Figs. 3 and 4, the results for FD seem to be quite the
same in serial and in parallel. For CS the accuracy seems higher in serial. For CD,
the accuracy is also higher in serial and, in serial, it has a ﬂuctuating behavior. Can
you comment on these observations?
{\color{red}  
\begin{itemize}
     \item
[TODO]
	The fluctuation seems to be a function of the amount of nonzero TS matrix elements relative to total TS matrix elements. Information supporting this has been added to the manuscript. This may be related to the fact that neighborhood radius in terms of number of Peridynamic nodes was kept constant and may indicate that connectivity in the mesh oscillates when the mesh is refined incrementally. 
  \end{itemize}}

23. Pgs. 17-18: What are the units of K ? If F int in Eq. (4) etc. is internal force, then K
should have units of force/length. However, in Figs. 3 and 4 you use the unit of
MPa, which is force/length2 . Please clarify.

24. Pg. 19: Are the units of the vertical axis in Fig. 5 “seconds”, as in Figs. 1 and 2?
Please indicate that. Also, is the label of the vertical axis in Fig. 6 correct?

25. Pg. 20: In lines 32-35, clarify what “byte-copyable” mean and why is that important.
{\color{red}  
\begin{itemize}
     \item
     Clarification made in manuscript, new reference added. 'Byte-copiable' means that members of an array of data elements are all of the same size in memory so that if one knew the position of the first element in the array, the position of all other elements in the array could be inferred with the additional information of the length of the array. This is important, in particular, to the designers and users of the thousands of MIC accelerator cards installed in the \emph{Stampede} computer system. MIC accelerator cards permit only the transfer of byte-copiable data from the host computer system for use with the '#pragma offload' mode, which is to say most code written for the MIC. This is even more important when it is realized that the great majority of \emph{Stampede}'s theoretical processing power is counted from these MIC accelerator cards. Meaning that if an AD user wanted to use any of that power, they would be out of luck since not even the full time personel at TACC could manage to compile the AD library Trilinos::Sacado for the MIC. A more 'casual' computational scientist would find the task impossible.
  \end{itemize}}

26. Pg. 21: Please include a section title for the references.

27. Pg. 24: In line 23, you may add the truncation error O(h2 ) to the expression. Oth-
erwise, you may need to replace the “=” sign by the “≈” sign. However, adding
the error term would make more clear the limiting expression when h → 0. Also,
replace the comma by a period.

28. Pg. 24: Why do we need S and X in Appendix B?
{\color{red}  
\begin{itemize}
     \item
     All the members of set X need to be members of set S otherwise the approximations don't work. X are numbers from the 'real' world and S are numbers from the computer world. For example: "http://en.wikipedia.org/wiki/IEEE_floating_point" shows that computers and languages following IEEE standards don't represent very large numbers with very fine precision at the same time due to memory constraints, and there are upper and lower bounds on the magnitudes of nonzero numbers that can be represented. Obviously, approximation error is expected, but the fact that a user doesn't actually get 'all' the numbers is something that is not obvious.
  \end{itemize}}

29. Pgs. 25-26: Are the functions in the second column of the table in 2(c) (page 25)
approximations? Why do you label the column “Approximates Function”? Simi-
larly, in line 16 (page 26) you say “approximated by the partial derivative”; is it an
approximation?
{\color{red}  
\begin{itemize}
     \item
      A computer can only provide finite precision approximations, of approxmations, of functions. Approximations of derivatives of functions are also approximations, therefor the label is correct in both cases. For example: "https://sourceware.org/git/?p=glibc.git;a=blob;f=sysdeps/x86_64/fpu/e_expl.S" is the implementation of the $exp(x)$ function on 80-bit extended precision floating point numbers in the popular GCC compiler. This implementation is the most accurate that GCC provides for the $e$ function on real numbers, but ultimately is still a discrete numerical algorithm.
  \end{itemize}}

30. Pg. 26: Please clarify whether storage is needed in automatic differentiation. For
example, do we need to store w, v , and u?

{\color{red}  
\begin{itemize}
     \item
      Clarification is added in the manuscript. For an actual AD system like Trilinos::Sacado or ADIFOR, w, v and u (stand-ins for simple or elementary math operations) need to be written in code differently from the built-in implementation to replace all math operations as part of a complete AD datatype. What doesn't absolutely have to be stored are the evaluations of w, v and u. Not storing the evaluations happens in 'forward' AD, storing the evaluations happens in 'backward' or 'reverse accumulation' AD.  
  \end{itemize}}

  \end{enumerate}
\end{document}

### 2014-02-27 jacobian_calc_CMAME.tex
more progress on II

The aim of this paper is not only to \note[Michael]{re: I.1}\change[Michael]{introduce the new}{study the} complex-step method \change[Michael]{in
the context of}{of} evaluating tangent-stiffness matrices, but to serve as a review
for other differentiation techniques useful for solving non-linear systems with
Newton's method. After presenting background information on the underlying
methods to be discussed, presented in this paper are: In subsection~\ref{subsec:TS} a description of
tangent-stiffness matrices, detailed directions for producing
tangent-stiffness matrices with each of the methods identified above, and a
description and justification of the \note[Michael]{re: I.1}\remove[Michael]{new}complex-step method for calculating
tangent-stiffness matrices, in sub-subsection~\ref{subsubsec:Impl} a description of implementing complex-step in
the \emph{Peridigm} software, in section~\ref{JGAM} a description of the quantities-of-interest
used to rank the methods, in section~\ref{sec:Res} a presentation and analysis of the results of the
comparative study, and in section~\ref{sec:Conc} conclusions and thoughts on potential future work.
Finally, in the interest of replicable research, the reader is referred to the
corresponding author's website for \emph{Peridigm} C++ source-code and data
necessary to reproduce the results presented here. Additionally, included on
the website is an example serial C++ library with classes for solving
non-linear systems using the techniques discussed here, as well as two
validation and verification example problems which show how to use the library.
The purpose of the serial C++ library is to demonstrate in a less complex
manner than in the \emph{Peridigm} code how one would use
automatic-differentiation or complex-step to better encourage their use and
discussion. This open source resource will be referred to as the \emph{paper
repository} in what follows.

@@

\subsubsection{The ``complex-step'' method.}  \label{sec:CSmethod} It is
possible to approximate derivatives quite accurately with a technique based on
a complex variable Taylor series expansion of a function.  This method was
first described by Lyness and Moler
\cite{lyness1967numerical,lyness1968differentiation} and has more recently been
rediscovered for use in engineering analysis.  The basic idea is that a model parameter can
be be made complex and expanded in a Taylor series about a small perturbation
along the imaginary axis by some arbitrary value $h$ as follows
%
\begin{align} f ( x + i h ) =& f(x) + \frac{\partial f(x)}{\partial x} \frac{i
h}{1!}  \notag \\ &+ \frac{\partial f^2(x)}{\partial x^2} \frac{(i h)^2}{2!} +
\frac{\partial f^3(x)}{\partial x^3} \frac{(i h)^3}{3!} + \cdots,
\label{eqn:complexTaylor} \end{align}
%
Taking the imaginary part of both sides of equation (\ref{eqn:complexTaylor})
and solving for the first derivative and ignoring terms of $\mathcal{O}\left (
h^2 \right)$ yields an estimate
%
\begin{equation} \frac{\partial f( x )}{\partial x} \approx \frac{\mbox{Im}
\left( f (x + i h) \right)}{h}.  \label{eqn:complexFirstDeriv} \end{equation}
%
Thus, an estimate of the first derivative of a function can be made by only
utilizing one functional evaluation of a perturbed model parameter along the
imaginary axis.  The step-size $h$ is arbitrary and can be made as small as
practical (even to machine precision without the dangers of roundoff error as
in finite-differences) to yield an accurate estimate of this derivative.  The
only disadvantage of this technique is the necessity of requiring the functions
to accept complex numbers as arguments. Appendix~\ref{sec:appendixA} includes a
simple example that illustrates the use of the complex-step method for the
derivative calculation of a function.

@@

An AD system evaluates composition functions in the expected manner: it works
by first evaluating the innermost function of the composition, then presenting
that function's output as input to the next level function until all levels are
complete, observing commonly expected order of operations. However, an AD
system does additional work during a function evaluation in that as each nested
function is evaluated, the function's partial derivative with respect to the
designated variables of the given input is also calculated.  This is possible
because the elementary math functions are hard-coded into the AD source-code
along with their analytical derivatives, and linked by special instructions, so
that when the elementary math functions are called upon for computation, their
partial derivatives may be computed and stored in a sequence. The AD system
then multiplies the final sequence of partial derivatives together to produce
the exact equivalent to taking a partial derivative of the corresponding
composition function with respect to a designated variable at a particular
value. It is obvious, but bears mentioning, that the AD system could simply
store one value for partial derivatives, modifying it as appropriate for every
function evaluation rather than keeping a sequence. This is significant because
modifying a single value rather than keeping a list of values corresponding to
every level of a composition function, which itself may not be the sole one
needing to be evaluated, represents a savings in memory usage, which at large
scales is an active concern. In the literature, the AD scheme described here is
called \emph{forward automatic-differentiation}. Only forward AD will be
covered here since it is the implementation used in \emph{Sacado} and therefore
in this study; however, the reader is referred to the introduction section of
\cite{ref-AD-methods} and its citations for further information on AD, and
particularly \cite{ref-on-AD} which is foundational.
Appendix~\ref{sec:appendixB} includes an example that walks the reader through
the process a computer uses to compute derivatives via AD.  Some things to note
about AD are that no approximation of derivatives is being made because the
analytical forms of the partials of the elementary math functions are defined
alongside them. The accuracy of AD is then limited by the precision of the AD
system's definition of the elementary math functions and their partial
derivatives.

@@

\subsection{Speed Data} Speed measurements were taken according to the methods
mentioned in Section~\ref{tcs}. The averaged results of these speed
measurements for the serial tests can be seen in Fig.~\ref{fig:serial_speed}.
Average time in seconds to compute a TS matrix is plotted as a function of the
number of nonzero TS elements.  CS was the slowest of the four methods, taking
the longest average time per iteration,  followed by CD, FD and AD in that
order. Calculation time has a power-law relationship with problem size, with a
power-law index of roughly $\num{1.0E-6}$ for each of the methods. This bodes
well for the scalability of each of the methods by themselves in
\emph{Peridigm}.
%
\begin{figure}[tbp] \centering \scalebox{1.0}{\input{./figs/serial_speed.pgf}}
\caption{Serial test series speed measurements.} \label{fig:serial_speed}
\end{figure}
%
For the parallel simulations, averaged speed results appear in
Fig.~\ref{fig:multi_speed}. For this series, rather than varying the number of
nonzero TS matrix elements, the number of compute cores used to solve the
problem is increased, while the number of nonzero TS matrix elements remained
constant at $\num{1.67E9}$ .  CS again was the slowest method, followed by CD,
AD and then FD.  There is a somewhat surprise reversal of FD and AD for these
parallel simulations.  Figure~\ref{fig:multi_speed} indicates that when looking
at average iteration time, the factor of speed improvement scales linearly with
the number of compute cores used for the given conditions of this group of
tests. Looking at Fig.~\ref{fig:multi_speed}, the reader can see that when core
count doubles, compute time halves, for each method. Again, these results show
that each of the methods as implemented in \emph{Peridigm} scale well when
additional processors are used. 
%
\begin{figure}[tbp] \centering \scalebox{1.0}{\input{./figs/multi_speed.pgf}}
\caption{Multicore test series speed measurements.} \label{fig:multi_speed}
\end{figure}
%
A speculative explanation as to why FD and AD have reversed places in the speed
tests with respect to the serial results is that the AD data-structures carry a
list of partial derivatives of the functional evaluations along side the
functional evaluations themselves only to be processed at the end of the
tangent-stiffness calculation when the derivatives are requested. This
represents additional load on the hardware memory subsystem, such that may
exceed that of FD yet still fall under a supposed memory bandwidth threshold
during serial operation. Cores within a multicore processor use the same shared
bus to access memory, such that cores compete for memory access. Supposing that
exceeding the memory bandwidth threshold penalizes additional requested
bandwidth, it is likely that the decrease in available memory bandwidth per
core caused by parallel operation coupled with AD's additional resource
requirements causes the AD method to be slower than FD.

>>>>

The aim of this paper is not only to \note[Michael]{re: I.1}\change[Michael]{introduce the new}{study the} complex-step method \change[Michael]{in
the context of}{of} evaluating tangent-stiffness matrices, but to serve as a review
for other differentiation techniques useful for solving non-linear systems with
Newton's method. After presenting background information on the underlying
methods to be discussed, presented in this paper are: In subsection~\ref{subsec:TS} a description of
tangent-stiffness matrices, detailed directions for producing
tangent-stiffness matrices with each of the methods identified above, and a
description and justification of the \note[Michael]{re: I.1}\remove[Michael]{new}complex-step method for calculating
tangent-stiffness matrices, in sub-subsection~\ref{subsubsec:Impl} a description of implementing complex-step in
the \emph{Peridigm} software, in section~\ref{JGAM} a description of the quantities-of-interest
used to rank the methods, in section~\ref{sec:Res} a presentation and analysis of the results of the
comparative study, and in section~\ref{sec:Conc} conclusions and thoughts on potential future work.
Finally, in the interest of replicable research, the reader is referred to the
corresponding author's website for \emph{Peridigm} C++ source-code and data
necessary to reproduce the results presented here. Additionally, included on
the website is an example serial C++ library with classes for solving
non-linear systems using the techniques discussed here, as well as two
validation and verification example problems which show how to use the library.
The purpose of the serial C++ library is to demonstrate in a less complex
manner than in the \emph{Peridigm} code how one would use
automatic-differentiation or complex-step to better encourage their use and
discussion. This open source resource is available at the correposinding author's website: http://idl.utsa.edu/jtfoster/paper-repositories/
 or directly at https://github.com/utsa-idl/Complex_Step_Jacobian_Project_Archive.

@@

\subsubsection{The ``complex-step'' method.}  \label{sec:CSmethod} It is
possible to approximate derivatives quite accurately with a technique based on
a complex variable Taylor series expansion of a function.  This method was
first described by Lyness and Moler
\cite{lyness1967numerical,lyness1968differentiation} and has more recently been
rediscovered for use in engineering analysis.  The basic idea is that a model parameter can
be be made complex and expanded in a Taylor series about a small perturbation
along the imaginary axis by some arbitrary value $h$ as follows
%
\begin{align} f ( x + i h ) =& f(x) + \frac{\partial f(x)}{\partial x} \frac{i
h}{1!}  \notag \\ &+ \frac{\partial^2 f(x)}{\partial x^2} \frac{(i h)^2}{2!} +
\frac{\partial^3 f(x)}{\partial x^3} \frac{(i h)^3}{3!} + \cdots,
\label{eqn:complexTaylor} \end{align}
%
Taking the imaginary part of both sides of equation (\ref{eqn:complexTaylor})
and solving for the first derivative and ignoring terms of $\mathcal{O}\left (
h^2 \right)$ yields an estimate
%
\begin{equation} \frac{\partial f( x )}{\partial x} = \frac{\mbox{Im}
\left( f (x + i h) \right)}{h} +\mathcal{O}\left ( h^2 \right).  \label{eqn:complexFirstDeriv} \end{equation}
%
Thus, an estimate of the first derivative of a function can be made by only
utilizing one functional evaluation of a perturbed model parameter along the
imaginary axis.  The step-size $h$ is arbitrary and can be made as small as
practical (even to machine precision without the dangers of roundoff error as
in finite-differences) to yield an accurate estimate of this derivative.  The
only disadvantage of this technique is the necessity of requiring the functions
to accept complex numbers as arguments. Appendix~\ref{sec:appendixA} includes a
simple example that illustrates the use of the complex-step method for the
derivative calculation of a function.

For the reader's convenience the foward difference equivalent to~\ref{eqn:complexFirstDeriv} is
%
\begin{equation} \frac{\partial f( x )}{\partial x} = \frac{f (x + h) - f (x)}{h} +\mathcal{O}\left ( h \right). \label{eqn:forwardFirstDeriv} \end{equation},
%
while the central difference equivalent to~\ref{eqn:complexFirstDeriv} is,
%
\begin{equation} \frac{\partial f( x )}{\partial x} = \frac{f (x + h) - f (x - h)}{2h} +\mathcal{O}\left ( h^2 \right). \label{eqn:centeredFirstDeriv} \end{equation} \cite[Chap. 4.1.3]{chapra2010}.
%
    Note that the order of the truncation error in~\ref{eqn:forwardFirstDeriv} indicates that for a given $h$ forward difference is less accurate an approximation than central difference or complex step.

@@

An AD system evaluates composition functions in the expected manner: it works
by first evaluating the innermost function of the composition, then presenting
that function's output as input to the next level function until all levels are
complete, observing commonly expected order of operations. However, an AD
system does additional work during a function evaluation in that as each nested
function is evaluated, the function's partial derivative with respect to the
designated variables of the given input is also calculated.  This is possible
because the elementary math functions are hard-coded into the AD source-code
along with their analytical derivatives, and linked by special instructions, so
that when the elementary math functions are called upon for computation, their
partial derivatives may be computed and stored in a sequence. The AD system
then multiplies the final sequence of partial derivatives together to produce
the exact equivalent to taking a partial derivative of the corresponding
composition function with respect to a designated variable at a particular
value. 
%
It is obvious, but bears mentioning, that the AD system could simply
store one value for partial derivatives, modifying it as appropriate for every
function evaluation rather than keeping a sequence. This is significant because
modifying a single value rather than keeping a list of values corresponding to
every level of a composition function, which itself may not be the sole one
needing to be evaluated, represents a savings in memory usage, which at large
scales is an active concern. In the literature, the AD scheme described here is
called \emph{forward automatic-differentiation}. Only forward AD will be
covered here since it is the implementation used in \emph{Sacado} and therefore
in this study; however, the reader is referred to the introduction section of
\cite{ref-AD-methods} and its citations for further information on AD, and
particularly \cite{ref-on-AD} which is foundational.
Appendix~\ref{sec:appendixB} includes an example that walks the reader through
the process a computer uses to compute derivatives via AD.  Some things to note
about AD are that no approximation of derivatives is being made because the
analytical forms of the partials of the elementary math functions are defined
alongside them. The accuracy of AD is then limited by the precision of the AD
system's definition of the elementary math functions and their partial
derivatives.

@@

\subsection{Speed Data} Speed measurements were taken according to the methods
mentioned in Section~\ref{tcs}. The averaged results of these speed
measurements for the serial tests can be seen in Fig.~\ref{fig:serial_speed}.
Average time in seconds to compute a TS matrix is plotted as a function of the
number of nonzero TS elements.  CS was the slowest of the four methods, taking
the longest average time per iteration,  followed by CD, FD and AD in that
order. Calculation time has a power-law relationship with problem size, with a
power-law index of roughly $\num{1.0E-6}$ for each of the methods. This bodes well
for the scalability of each of the methods by themselves in \emph{Peridigm}. If we
assumed every individual operation, generally speaking, took the same amount of
time on average such that compute time were a measure of number of operations,
this suggests that the methods as in \emph{Peridigm} have the same order
algorithm complexity when considering computing a TS matrix alone, that is
computation time is $\mathcal{O}(10^{(10^{-6} \times \log_{10}n)})$. This would
suggest, to the algorithm design way of thinking, that the methods are equally
well suited to the task of computing a TS matrix, if it were further assumed
that the TS matrices produced by each method were identical. This is not the case
as seen in the following subsection.
%
\begin{figure}[tbp] \centering \scalebox{1.0}{\input{./figs/serial_speed.pgf}}
\caption{Serial test series speed measurements.} \label{fig:serial_speed}
\end{figure}
%
For the parallel simulations, averaged speed results appear in
Fig.~\ref{fig:multi_speed}. For this series, rather than varying the number of
nonzero TS matrix elements, the number of compute cores used to solve the
problem is increased, while the number of nonzero TS matrix elements remained
constant at $\num{1.67E9}$ .  CS again was the slowest method, followed by CD,
AD and then FD.  There is a somewhat surprise reversal of FD and AD for these
parallel simulations.  Figure~\ref{fig:multi_speed} indicates that when looking
at average iteration time, the factor of speed improvement scales inversely with
the number of compute cores used for the given conditions of this group of
tests. Looking at Fig.~\ref{fig:multi_speed}, the reader can see that when core
count doubles, compute time halves, for each method. Again, these results show
that each of the methods as implemented in \emph{Peridigm} scale well when
additional processors are used. 
%
\begin{figure}[tbp] \centering \scalebox{1.0}{\input{./figs/multi_speed.pgf}}
\caption{Multicore test series speed measurements.} \label{fig:multi_speed}
\end{figure}
%
A speculative explanation as to why FD and AD have reversed places in the speed
tests with respect to the serial results is that the AD data-structures carry a
list of partial derivatives of the functional evaluations along side the
functional evaluations themselves only to be processed at the end of the
tangent-stiffness calculation when the derivatives are requested. This
represents additional load on the hardware memory subsystem, such that may
exceed that of FD yet still fall under a supposed memory bandwidth threshold
during serial operation. Cores within a multicore processor use the same shared
bus to access memory, such that cores compete for memory access. Supposing that
exceeding the memory bandwidth threshold penalizes additional requested
bandwidth, it is likely that the decrease in available memory bandwidth per
core caused by parallel operation coupled with AD's additional resource
requirements causes the AD method to be slower than FD.
